The \project{unRAVE} catalog

### 2016-08-26 article/unrave.tex
new problem..

\newcommand{\argmin}[1]{\underset{#1}{\operatorname{argmin}}\,}

@@

The first \project{Gaia} data release will include positions, proper motions, and 
parallaxes for approximately two million stars in the Tycho-2 \citep{Hog_2000} 
catalog \citep{Michalik_2015a,Michalik_2015b}.  After cross-matching all major 
stellar spectroscopic surveys\footnote{Specifically we cross-matched the Tycho-2
catalog against the \apogee\ \citep{Zasowski_2013}, \ges\
\citep{Gilmore_2012,Randich_2013}, \galah\ \citep{DeSilva_2015},
\lamost\ \citep{Cui_2012}, and \rave\ \citep{Steinmetz_2006} 
surveys}, we found the RAdial Velocity Experiment (hereafter \rave) survey to 
have the largest overlap with the first \gaia\ data release: 292,036 
stars.  We then used the \gaia\ universe model snapshot \citep{Robin_2012} to 
estimate the precision in parallax and proper motions that will be available in 
the first \gaia\ data release (DR1) for stars in those overlap samples. Comparing
the expected precision to what is currently available, we further found that the
\rave\ survey will benefit most from \gaia\ DR1.  The distances of 63\% of the 
\rave--\gaia\ DR1 overlap sample (182,862 stars) are expected to improve with 
the first \gaia\ data release, and 47\% of stars are likely to have better 
proper motions (137,211 stars).  This motivated us to examine what chemical 
abundance information was available for \rave, and to evaluate whether we could
enable new chemodynamic studies by contributing to this existing set of chemical
abundances.

@@

We have constructed a justified training set for stars across the main-sequence, sub-giant,
and red giant branch.  However the lack of overlap between \rave\ and other works have
resulted in a somewhat peculiar situation. Detailed abundances are available from \apogee\
for all giant stars in our sample, but only imprecise metallicities are available from
\epic\ for stars on the main-sequence and the sub-giant branch.  In the following section
we will construct a simplified model that does not include detailed abundances, before
demonstrating how we (1) improve upon the imprecise metallicities for main-sequence/sub-giant
stars, and (2) translate abundances from the giant branch stars to the main-sequence/sub-giant
branch sample.

@@

% Describe The Cannon basic model, vectorizer, etc.
% Fitting time errors, etc.

% Here we will first consider only a three parameter model.
% APOGEE used only quadratic model, worked for giants
% We experimented with a quadratic model only. 
% Useful for identifying mis-classifications but the structure wasn't all there..
% worked but didn't show the structure in detail.
% Split into two different models with partial overlap.
% One for giants and one for dwarfs.
% The labels for stars in the training sets are shown in Figure X.

% We trained all three models.
% Note the differences in metallicity scales.
% The K2/EPIC stuff is accurate but imprecise, whereas the APOGEE stuff is precise, usually precise (except Ti?)
% Look at spectral derivatives.
% Show the model internals are relatable.
% Map spectral derivatives from giant model to main-sequence model.
% Update the model with those new spectral derivatives.

% We ran all three models on all spectra.
% How we combined the results into one.
% Refer to test set HRD.

% Restricted to 3-label model, but the giant set includes abundance labels.
% The spectral range covered by \rave\ includes atomic transitions of X, Y, Z elements.
% The strongest lines are indicated in Figure XXX.

% Censoring masks
% Transfering coefficients, etc.

>>>>

\newcommand{\Dvector}[1]{\boldsymbol{#1}}
\newcommand{\vectheta}{\Dvector{\theta}}
\newcommand{\vecv}{\Dvector{v}}
\newcommand{\argmin}[1]{\underset{#1}{\operatorname{argmin}}\,}

@@

The first \project{Gaia} data release will include positions, proper motions, and 
parallaxes for approximately two million stars in the Tycho-2 \citep{Hog_2000} 
catalog \citep{Michalik_2015a,Michalik_2015b}.  After cross-matching all major 
stellar spectroscopic surveys\footnote{Specifically we cross-matched the Tycho-2
catalog against the \apogee\ \citep{Zasowski_2013}, \ges\
\citep{Gilmore_2012,Randich_2013}, \galah\ \citep{DeSilva_2015},
\lamost\ \citep{Cui_2012}, and \rave\ \citep{Steinmetz_2006} 
surveys}, we found the RAdial Velocity Experiment (hereafter \rave) survey to 
have the largest overlap with the first \gaia\ data release: 292,036 
stars.  We then used the \gaia\ universe model snapshot \citep{Robin_2012} to 
estimate the precision in parallax and proper motions that will be available in 
the first \gaia\ data release (DR1) for stars in those overlap samples. Comparing
the expected precision to what is currently available, we further found that the
\rave\ survey will benefit most from \gaia\ DR1.  The distances of 63\% of the 
\rave--\gaia\ DR1 overlap sample (182,862 stars) are expected to improve with 
the first \gaia\ data release, and 47\% of stars are likely to have better 
proper motions (137,211 stars).  This motivated us to examine what chemical 
abundances were available for \rave, and to evaluate whether we could enable new 
chemodynamic studies by contributing to this existing set of chemical abundances.

@@

We have constructed a justified training set for stars across the main-sequence, sub-giant,
and red giant branch.  However the lack of overlap between \rave\ and other works have
resulted in a somewhat peculiar situation. Detailed abundances are available from \apogee\
for all giant stars in our sample, but only imprecise metallicities are available from
\epic\ for stars on the main-sequence and the sub-giant branch.  To our knowledge this 
problem has not been addressed with data-driven models in the astrophysics literature.
In the following section we will construct a simplified model that does not include detailed
abundances, before demonstrating how we (1) improve upon the imprecise metallicities for 
main-sequence/sub-giant stars, and (2) translate abundances from the giant branch stars 
to the main-sequence/sub-giant branch sample.

@@

\noindent{}The model we adopt is
\begin{eqnarray}\label{eq:model}
y_{jn} & = & \vecv(\ell_n)\cdot\vectheta_j + e_{jn}\quad ,
\end{eqnarray}

\noindent{}where $y_{jn}$ is the pseudo-continuum-normalized flux for star $n$ at wavelength pixel
$j$, $\vecv(\ell_n)$ is the vectorizing function that takes as input the $K$ labels
$\ell_n$ for star $n$ and outputs functions of those labels as a vector of length
$D>K$, $\vectheta_j$ is a vector of length $D$ of parameters influencing the model at
wavelength pixel $j$, and $e_{jn}$ is the residual (noise).  Here we will only consider
vectorizing functions with polynomial expansions (e.g., $\teff^2$, see Section 
\ref{sec:a-simple-model}).  The noise values $e_{jn}$ can be considered to be drawn 
from a Gaussian distribution with zero mean and variance $\sigma_{jn}^2 + s_j^2$, where
$\sigma_{jn}^2$ is the variance in flux $y_{jn}$ and $s_j^2$ describes the excess 
variance at the $j$-th wavelength pixel.  We direct the reader to \citet{Ness_2015} for
a more detailed introduction and justification of \thecannon.

At the training step we fix the $K$-lists of labels for all the training set stars $n$.
At each wavelength pixel $j$, we then calculate the parameters $\vectheta_j$ and $s_j^2$
by optimizing the likelihood function:
\begin{eqnarray}\label{eq:train}
\vectheta_j,s^2_j &\leftarrow& \argmin{\vectheta,s}\left[
    \sum_{n=0}^{N-1} \frac{[y_{jn}-\vecv(\ell_n)\cdot\vectheta]^2}{\sigma^2_{jn}+s^2}
    + \sum_{n=0}^{N-1} \ln(\sigma^2_{jn}+s^2)
    \right]
  \quad .
\end{eqnarray}

In practice this is achieved by fixing $s_j^2 = 0$ to make equation~\ref{eq:train} a
convex optimization problem, then we optimize for $\vectheta_j$, before solving for $s_j^2$.
We can also adopt strict priors on $\vectheta$ such that labels can only make non-zero
contributions to the data in certain wavelength ranges.  That is to say, we can specify
that certain labels (e.g., [Al/H]) can only contribute to the model at wavelength pixels
where we know an aluminium transition exists.  This method, which we call \emph{censoring},
is described in Section~\ref{sec:censoring} and \citet{Ness_2016b}.

The \emph{test step} is where we fix the parameters $\vectheta_j,s_j^2$ at all wavelength
pixels $j$, and optimize the $K$-list of labels $\ell_m$ for each test set star $m$.  Here
the likelihood function is:
\begin{eqnarray}\label{eq:test}
  \ell_m &\leftarrow& \argmin{\ell}\left[
    \sum_{j=0}^{J-1} \frac{[y_{jm}-\vecv(\ell)\cdot\vectheta_j]^2}{\sigma_{jm}^2 + s_j^2}
    \right]
  \quad .
\end{eqnarray}

At the test step optimization we store the covariance matrix $\bm{\Sigma}$ for the labels
$\ell$, which provides us with the formal errors on $\ell$.  We judge the veracity of these
formal errors in our validation experiments in Section \ref{sec:validation}.

\subsection{Three-label models ($\teff$, $\logg$, [Fe/H])}
\label{sec:a-simple-model}

% Given previous successes, we will first adopt a three-parameter model for giants + dwarfs.
% Polynomial thing --> expand out \vectheta
% worked but didn't show the structure in detail.
% Split into two different models with partial overlap.
% One for giants and one for dwarfs.
% The labels for stars in the training sets are shown in Figure X.

% We trained all three models.
% Note the differences in metallicity scales.
% Look at spectral derivatives.
% Show the model internals are relatable.
% Map spectral derivatives from giant model to main-sequence model.
% Update the model with those new spectral derivatives.

% How we combined the results into one.
% Refer to test set HRD.

\subsection{A 10-label model for detailed abundances of giant stars}
\label{sec:giant-abundance-model}
% Restricted to 3-label model, but the giant set includes abundance labels.
% The spectral range covered by \rave\ includes atomic transitions of X, Y, Z elements.
% The strongest lines are indicated in Figure XXX.
% Censoring masks

% Train
